---
title: "Building Synthetic Risk-Analysis Fixtures"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Building Synthetic Risk-Analysis Fixtures}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

# Overview

`riskGenerateR` generates reproducible synthetic spatial and contextual data
for developing and testing risk-related analytical workflows.

The package follows a deliberate separation of responsibilities:

```text
spatial geometry
      |
      v
contextual attributes
      |
      v
synthetic analytical fixture
      |
      v
downstream analysis
```

`riskGenerateR` generates the fixture.

It does not calculate or infer analytical risk outputs such as risk scores,
risk classes, rates, expected counts or standardised ratios.

This makes the generated data suitable for testing different analytical
methods without embedding the analytical answer in the synthetic input.

# Load the package

```{r load-package}
library(riskGenerateR)
```

# Build a common spatial environment

A useful feature of the package is that geometry and semantic context are
separate.

Start with a neutral synthetic study area and a set of points.

```{r spatial-environment}
study_area <- rgr_study_area(
  bbox = c(
    xmin = 0,
    ymin = 0,
    xmax = 5000,
    ymax = 4000
  ),
  area_id = "fixture_area",
  area_name = "Synthetic Fixture Area"
)

points <- rgr_points(
  study_area,
  n = 100,
  inside_pct = 1,
  seed = 123
)

points
```

At this stage the points contain spatial-generation information, but they do
not yet represent incidents, hazards, observations or other risk-related
records.

```{r point-fields}
names(
  points
)
```

# Add contextual attributes

`rgr_add_attributes()` assigns a synthetic semantic context to existing
spatial features.

The currently supported contexts are:

- `hazard_assessment`
- `incident`
- `risk_register`
- `observation`
- `generic`

The geometry is retained while contextual fields are added.

# Hazard-assessment fixtures

A hazard-assessment fixture represents plausible assessment inputs.

```{r hazard-assessment}
hazards <- rgr_add_attributes(
  points,
  type = "hazard_assessment",
  seed = 101
)

names(
  hazards
)
```

The generated records include contextual fields describing matters such as
the hazard, source, exposed group, existing controls, consequence and
likelihood.

```{r hazard-example}
head(
  sf::st_drop_geometry(hazards)
)
```

These are synthetic assessment inputs.

They are not converted by `riskGenerateR` into a calculated risk score or
risk class.

```{r hazard-summary}
rgr_summary(
  hazards
)
```

# Incident fixtures

The same point geometry can instead represent synthetic incidents.

```{r incident}
incidents <- rgr_add_attributes(
  points,
  type = "incident",
  seed = 102
)

names(
  incidents
)
```

Incident records include contextual variables such as event type, hazard
category, mechanism, activity, consequence and event time.

```{r incident-example}
head(
  sf::st_drop_geometry(incidents)
)
```

An `event_count` field is included as a descriptive count associated with
each generated event record.

```{r incident-summary}
rgr_summary(
  incidents
)
```

# Risk-register fixtures

Synthetic risk-register records can be generated from the same geometry.

```{r risk-register}
risk_register <- rgr_add_attributes(
  points,
  type = "risk_register",
  seed = 103
)

names(
  risk_register
)
```

These records describe plausible risk-management information including
objectives, risk categories, causes, risk events, consequences, existing
controls, ownership and review status.

```{r risk-register-example}
head(
  sf::st_drop_geometry(risk_register)
)
```

The presence of risk-management terminology does not mean that the package
has calculated risk.

The generated records provide inputs that downstream workflows can assess,
classify or analyse.

```{r risk-register-summary}
rgr_summary(
  risk_register
)
```

# Observation fixtures

Observation records provide a lighter descriptive context suitable for field,
inspection or monitoring-style workflows.

```{r observation}
observations <- rgr_add_attributes(
  points,
  type = "observation",
  seed = 104
)

names(
  observations
)
```

Generated fields describe the observation, condition, issue, descriptive
severity indicator, required action, action status and observation date.

```{r observation-example}
head(
  sf::st_drop_geometry(observations)
)
```

```{r observation-summary}
rgr_summary(
  observations
)
```

# Generic analytical fixtures

Not every test requires domain-specific terminology.

The `generic` context provides neutral synthetic variables that can be used
for general analytical development.

```{r generic}
generic <- rgr_add_attributes(
  points,
  type = "generic",
  seed = 105
)

names(
  generic
)
```

```{r generic-example}
head(
  sf::st_drop_geometry(generic)
)
```

This is useful when testing grouping, filtering, summarisation or other
general-purpose workflows without imposing a particular risk-management
meaning.

```{r generic-summary}
rgr_summary(
  generic
)
```

# Geometry does not determine meaning

The attribute context is independent of geometry type.

For example, create a square tessellation:

```{r grid}
grid <- rgr_tessellate(
  study_area,
  cell_size = 500,
  shape = "square",
  clip = TRUE
)
```

The cells can then receive generic contextual attributes.

```{r grid-attributes}
grid_attributes <- rgr_add_attributes(
  grid,
  type = "generic",
  seed = 201
)

rgr_summary(
  grid_attributes
)
```

This separation allows the same geometry-generation tools to support many
different testing scenarios.

# Generate a complete scenario

For common development and demonstration workflows, `rgr_scenario()` combines
the spatial and contextual generators into a reproducible fixture.

```{r basic-scenario}
scenario <- rgr_scenario(
  scenario = "basic",
  bbox_name = "wa_outback",
  n_points = 100,
  inside_pct = 0.9,
  cell_size = 500,
  seed = 123
)

names(
  scenario
)
```

The basic scenario provides a study area, generated spatial features and
contextual versions of those features.

The incident fixture can be inspected directly:

```{r scenario-incidents}
rgr_summary(
  scenario$incidents
)
```

Likewise, contextual attributes associated with the tessellated spatial units
can be inspected independently.

```{r scenario-grid}
rgr_summary(
  scenario$grid_attributes
)
```

```{r scenario-hex}
rgr_summary(
  scenario$hex_attributes
)
```

# Alternative scenario structures

`rgr_scenario()` can also generate more focused spatial configurations.

A square-grid scenario:

```{r square-scenario}
square_scenario <- rgr_scenario(
  scenario = "square_grid",
  seed = 123
)

names(
  square_scenario
)
```

A hexagonal-grid scenario:

```{r hex-scenario}
hex_scenario <- rgr_scenario(
  scenario = "hex_grid",
  seed = 123
)

names(
  hex_scenario
)
```

And an irregular-polygon scenario:

```{r polygon-scenario}
polygon_scenario <- rgr_scenario(
  scenario = "irregular_polygons",
  seed = 123
)

names(
  polygon_scenario
)
```

These provide reproducible starting points for testing workflows that require
different forms of spatial support.

# Reproducibility

Synthetic fixtures are often most useful when a failing test or analytical
example can be recreated exactly.

Supplying the same seed produces the same scenario.

```{r reproducibility}
scenario_again <- rgr_scenario(
  scenario = "basic",
  bbox_name = "wa_outback",
  n_points = 100,
  inside_pct = 0.9,
  cell_size = 500,
  seed = 123
)

identical(
  sf::st_as_binary(sf::st_geometry(scenario$points)),
  sf::st_as_binary(sf::st_geometry(scenario_again$points))
)
```

Contextual generation is also reproducible when the same seed and input are
used.

# Inspect before analysing

`rgr_summary()` provides a compact QA and descriptive view of generated
objects.

```{r summary-example}
rgr_summary(
  scenario$incidents
)
```

Its role is inspection rather than analytical inference.

Depending on the generated context, it can report:

- row and column counts
- geometry types
- CRS information
- geometry validity
- empty geometries
- attribute-context counts
- descriptive event or generic counts
- area summaries
- contextual category distributions

It intentionally does not calculate risk metrics.

# Package boundary

A typical workflow can therefore be thought of as:

```text
riskGenerateR
    |
    +-- generate study area
    +-- generate spatial features
    +-- generate contextual attributes
    +-- compose reproducible scenarios
    +-- perform fixture QA
    |
    v
synthetic analytical fixture
    |
    v
downstream package or application
    |
    +-- spatial joins
    +-- aggregation
    +-- rates
    +-- statistical modelling
    +-- risk analysis
    +-- decision support
```

This boundary is intentional.

It allows `riskGenerateR` to remain a reusable synthetic-data generator while
different downstream tools determine how those data should be analysed.

# Summary

`riskGenerateR` provides two complementary layers:

1. spatial generation through functions such as `rgr_study_area()`,
   `rgr_points()`, `rgr_tessellate()` and `rgr_polygons()`; and
2. contextual generation through `rgr_add_attributes()` and
   `rgr_scenario()`.

Together these functions create reproducible synthetic fixtures for
risk-related software development, testing, demonstrations and QA.

The package generates plausible inputs.

The analytical interpretation remains downstream.