---
title: "Getting Started with Spatial Generation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with Spatial Generation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

# Overview

`riskGenerateR` generates reproducible synthetic spatial and contextual data
for risk-related development, testing, demonstrations and quality assurance.

The spatial generation workflow is intentionally modular:

```text
study area
    |
    +-- points
    |
    +-- square or hexagonal tessellations
    |
    +-- irregular polygons
```

These spatial features can be used directly as synthetic test data or passed
to contextual attribute generators elsewhere in the package.

`riskGenerateR` generates synthetic data. It does not calculate or infer risk.

# Load the package

```{r load-package}
library(riskGenerateR)
```

The examples below use `sf` objects returned directly by `riskGenerateR`.

# Create a study area

A study area provides the spatial extent used by the geometry generators.

The simplest workflow uses a named numeric bounding box.

```{r study-area-xy}
study_area_xy <- rgr_study_area(
  bbox = c(
    xmin = 0,
    ymin = 0,
    xmax = 5000,
    ymax = 4000
  ),
  area_id = "example_xy",
  area_name = "Synthetic XY Study Area"
)

study_area_xy
```

This creates a synthetic coordinate space without assigning a coordinate
reference system.

Synthetic XY coordinates are useful when the purpose of the fixture is to
test geometry operations, package behaviour or analytical workflows without
representing a real location.

```{r study-area-xy-summary}
rgr_summary(
  study_area_xy
)
```

# Generate random points

`rgr_points()` generates random point features relative to a study area.

```{r points}
points <- rgr_points(
  study_area_xy,
  n = 100,
  inside_pct = 0.9,
  outside_distance = 500,
  seed = 123
)

points
```

The `inside_pct` argument controls the proportion generated inside the study
area.

Here, 90% of points are generated inside the study area and the remainder are
generated outside it.

```{r points-table}
table(
  points$inside_flag
)
```

A seed can be supplied when reproducible output is required.

```{r points-reproducibility}
points_again <- rgr_points(
  study_area_xy,
  n = 100,
  inside_pct = 0.9,
  outside_distance = 500,
  seed = 123
)

identical(
  sf::st_as_binary(sf::st_geometry(points)),
  sf::st_as_binary(sf::st_geometry(points_again))
)
```

# Generate a square tessellation

Regular spatial units can be generated with `rgr_tessellate()`.

```{r square-grid}
square_grid <- rgr_tessellate(
  study_area_xy,
  cell_size = 500,
  shape = "square",
  clip = FALSE
)

square_grid
```

With `clip = FALSE`, cells intersecting the study area are retained in their
regular form.

```{r square-summary}
rgr_summary(
  square_grid
)
```

To clip cells to the study-area boundary, use:

```{r square-grid-clipped}
square_grid_clipped <- rgr_tessellate(
  study_area_xy,
  cell_size = 500,
  shape = "square",
  clip = TRUE
)

square_grid_clipped
```

# Generate a hexagonal tessellation

The same function generates hexagonal cells.

```{r hex-grid}
hex_grid <- rgr_tessellate(
  study_area_xy,
  cell_size = 500,
  shape = "hex",
  clip = FALSE
)

hex_grid
```

Square and hexagonal tessellations use the same output structure, making it
straightforward to substitute one spatial support for another during
development and testing.

# Generate irregular polygons

`rgr_polygons()` generates irregular polygon units within a study area.

```{r irregular-polygons}
polygons <- rgr_polygons(
  study_area_xy,
  target_n = 20,
  cell_size = 100,
  seed = 123
)

polygons
```

The generated polygons provide less regular spatial units for workflows where
a square or hexagonal tessellation would be too artificial.

```{r polygon-summary}
rgr_summary(
  polygons
)
```

# Visual inspection

The generated geometries can be inspected with standard `sf` plotting.

```{r plot-points, fig.cap="Synthetic study area and generated points."}
plot(
  sf::st_geometry(study_area_xy),
  main = "Synthetic points"
)

plot(
  sf::st_geometry(points),
  add = TRUE
)
```

```{r plot-square, fig.cap="Synthetic square tessellation."}
plot(
  sf::st_geometry(square_grid),
  main = "Square tessellation"
)

plot(
  sf::st_geometry(study_area_xy),
  add = TRUE,
  lwd = 2
)
```

```{r plot-hex, fig.cap="Synthetic hexagonal tessellation."}
plot(
  sf::st_geometry(hex_grid),
  main = "Hexagonal tessellation"
)

plot(
  sf::st_geometry(study_area_xy),
  add = TRUE,
  lwd = 2
)
```

```{r plot-polygons, fig.cap="Synthetic irregular polygon units."}
plot(
  sf::st_geometry(polygons),
  main = "Irregular polygons"
)
```

# CRS-aware generation

Synthetic XY coordinates are not the only supported workflow.

`riskGenerateR` can also generate features using a defined coordinate
reference system.

The package includes example bounding boxes that can be used for reproducible
examples.

```{r bbox-example}
ex <- rgr_bbox_example(
  "wa_outback"
)

ex
```

The example can be passed directly to `rgr_study_area()`.

```{r projected-study-area}
study_area_projected <- rgr_study_area(
  bbox = ex$bbox,
  crs = ex$crs,
  area_id = "wa_outback_example",
  area_name = "WA Outback Example"
)

study_area_projected
```

Spatial features generated from this study area retain its spatial reference.

```{r projected-points}
projected_points <- rgr_points(
  study_area_projected,
  n = 50,
  seed = 456
)

sf::st_crs(
  projected_points
)
```

The same principle applies to tessellations and irregular polygons.

# Geometry generation and semantic context

Geometry generation is deliberately separate from the meaning attached to
the generated features.

For example, these points are simply synthetic point geometries:

```{r geometry-only}
names(
  points
)
```

They can subsequently be given a contextual meaning with
`rgr_add_attributes()`.

```{r incident-context}
incidents <- rgr_add_attributes(
  points,
  type = "incident",
  seed = 456
)

names(
  incidents
)
```

The geometry has not changed. The additional fields describe a synthetic
incident context.

This separation is central to the package design:

```text
geometry
    |
    v
contextual attributes
    |
    v
synthetic analytical fixture
    |
    v
downstream analysis
```

The contextual attribute workflow is covered in the
**Building Synthetic Risk-Analysis Fixtures** vignette.

# Summary

The core spatial generators in `riskGenerateR` are:

- `rgr_study_area()` for spatial extents
- `rgr_points()` for random point features
- `rgr_tessellate()` for square and hexagonal units
- `rgr_polygons()` for irregular polygon units

Together they provide reproducible spatial building blocks for testing,
demonstrations, QA and downstream risk-related analytical development.

Spatial generation remains separate from analytical risk calculation.